---
title: "data_wrangling_i"
output: github_document
date: "2023-09-19"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(readxl)
library(haven)
```
Importing the `FAS_litters.csv` csv. 
```{r}

litters_df= 
    read_csv("data/data_import_examples/FAS_litters.csv")

#makes everything lowercase, removes special characters, removes spaces
litters_df= 
  janitor::clean_names(litters_df)

litters_df

```
Importing the `FAS_pups.csv` csv. 
```{r}

pups_df= 
  read_csv("data/data_import_examples/FAS_pups.csv") %>%
  janitor::clean_names()

pups_df
```
## Look at data

```{r, eval=FALSE}
litters_df

head(litters_df)

tail(litters_df)
```

```{r}
str(litters_df)

skimr::skim(litters_df)

```
## Options in *read_*

The read function has some fun tools to help: 
  - skip= X will remove the first X rows and use the next as variable names 
  - colnames = FALSE will not to use the first row as varibale names, will be numbered columns instead
  - na = c(NA, "19", ".") shows missing values are equal to NA or "19" or "."-> replaces 19 with NA
  - coltypes= cols(Group= col_factor())
```{r}
litters_df_edited= 
  read_csv("data/data_import_examples/FAS_litters.csv",
           na=c("NA", 19, "."), 
           col_types = cols(Group= col_factor(), `GD0 weight`=col_character())
           ) 

```

## Other file types

Importing excel files
*read_excel* functions very similarly to *read_csv*. It's tidyverse adjacent, just not part of the tidyverse package. 
```{r}
mlb_df= 
  read_excel("data/data_import_examples/mlb11.xlsx")

mlb_df
```

Importing SAS files 
use *haven* package for SAS files 
```{r}
pulse_df= 
  read_sas("data/data_import_examples/public_pulse_data.sas7bdat")

pulse_df
```





